---
title: "EDA for nfl_yards"
author: "Kory D. Johnson"
date: 'Last Compiled `r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: show
    highlight: tango
    number_sections: yes
    theme: cosmo
    toc: yes
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE,
               collapse = FALSE,
               comment = "",
               strip.white = TRUE,
               warning = FALSE,
               message = FALSE,
               out.width = "70%",
               fig.align = "center")
```

## Topic

Defining features. Here because faster to iterate than python.

## Set Up  {.tabset .tabset-fade}

### Packages/Files {-}

```{r}
library(tidyverse)
library(magrittr)
```

### Load Data {-}

```{r, cache=T}
loadData <- function(fileName = "train.csv", ...) {
  file = paste("../input", fileName, sep="/")
  data.table::fread(file, ...) %>% 
    as_tibble()
}
df = loadData("trainClean.csv")
```

# Features w/o Players

## Distance to Goal

```{r}
df = df %>% 
  mutate(YardsToGoal = YardLine + (100-2*YardLine)*(PossessionTeam==FieldPosition))
df %>% 
  distinct(PlayId, .keep_all=T) %>% 
  ggplot(aes(x=YardsToGoal)) +
  geom_histogram()
```

Why is 75 ish so much more common?

## Line of Scrimmage
```{r}
df %<>% 
  mutate(LOS_Side = ifelse((PlayDirection=="right" & PossessionTeam==FieldPosition) |
                             (PlayDirection=="left" & PossessionTeam!=FieldPosition),
                           "left", "right"),
         LineOfScrimmage = ifelse(LOS_Side=="left", YardLine+10, 110-YardLine))
df %>% 
  distinct(PlayId, .keep_all = T) %>% 
  ggplot(aes(LineOfScrimmage)) +
  geom_histogram()
dfLOS = df %>% 
  group_by(PlayId) %>% 
  summarise(DistanceToLOS = mean(abs(X-LineOfScrimmage))) %>% 
  arrange(desc(DistanceToLOS))
ggplot(dfLOS) +
  geom_histogram(aes(DistanceToLOS))
longDist = dfLOS %>% 
  pull(PlayId)
index=1
dfPlay = filter(df, PlayId==longDist[index])
plotPlay(dfPlay, "DistanceToLOS", dfLOS[1, "DistanceToLOS"])
dfPlay %>% 
  filter(NflId==NflIdRusher) %>% 
  select(Dir, Orientation, S)
```

#  Checking python data
```{r}
dfPy = loadData("trainClean_py.csv") 
newFeatures = colnames(dfPy)[33:ncol(dfPy)]
for (col in newFeatures) {
  p = ggplot(dfPy, aes(!!sym(col))) +
    geom_density()
  print(p)
}
dfPy %>% 
  ggplot(aes(LineOfScrimmage)) +
  geom_histogram()
dfPy %>% 
  ggplot(aes(DistanceToGoal)) +
  geom_histogram()
dfPy %>% 
  ggplot(aes(DistanceToLOS)) +
  geom_histogram()
ggplot(dfPy) +
  geom_histogram(aes(DistanceToLOS))
longDist = dfPy %>% 
  pull(PlayId) %>% 
  head()
index=1
plotPlay(filter(df, PlayId==longDist[index]), "DistanceToLOS", dfPy[1, "DistanceToLOS"])
dfPy %>% 
  filter(PlayId==longDist[index])
```

# Features with Players

test df
```{r}
dfPlay = filter(df, PlayId == df$PlayId[1])
dfPlay
plotPlay(dfPlay)
```

## Distance to Defender

find runner; get opposing team, min l2 dist

determine offense by who **actually runs**; see play with largest distance to LOS

```{r}
runner = dfPlay %>% 
  filter(NflIdRusher == NflId) %>% 
  select(X, Y, Team)
blocker = dfPlay %>% 
  filter(Team != runner$Team) %>% 
  mutate(dist = sqrt((X-pull(runner,X))^2 + (Y-pull(runner,Y))^2)) %>% 
  summarise(
    dist = min(dist),
    blockerId = NflId[which.min(dist)]
  )
plotPlay(dfPlay)

```

