---
title: "Notes on "
author: "Kory D. Johnson"
date: 'Last Compiled `r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: show
    highlight: tango
    number_sections: yes
    theme: cosmo
    toc: yes
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE,
               collapse = FALSE,
               comment = "",
               strip.white = TRUE,
               warning = FALSE,
               message = FALSE,
               out.width = "70%",
               fig.align = "center")
```

## Topic

This file 

## Set Up  {.tabset .tabset-fade}

### Packages/Files {-}

```{r}
library(tidyverse)
library(magrittr)
```

### Load Data {-}

```{r, cache=T}
loadData <- function(fileName = "train.csv", ...) {
  file = paste("../input", fileName, sep="/")
  data.table::fread(file, ...) %>% 
    as_tibble()
}
df = loadData("trainClean.csv")
```

# Plotting Function

line of scrimmage isn't necessarily YardsToGoal, depnds on the side and direction of play.
If PlayDirection == right and in opponents field, then LOS is on right
If PlayDirection == right and in your field, then LOS is on Left
If PlayDirection == left and in opponents field, then LOS is on Left
If PlayDirection == leftand in your field, then LOS is on right.
Need to either add 10 or subtract 10 for endzone...

```{r}
dfPlay = filter(df, PlayId == df$PlayId[1]) %>% 
  mutate(YardsToGoal = YardLine + (100-2*YardLine)*(PossessionTeam==FieldPosition),
         LOS_Side = ifelse((PlayDirection=="right" & PossessionTeam==FieldPosition) |
                             (PlayDirection=="left" & PossessionTeam!=FieldPosition),
                           "left", "right"),
         LineOfScrimmage = ifelse(LOS_Side=="left", YardLine+10, 110-YardLine))
select(dfPlay, YardLine, YardsToGoal, LineOfScrimmage)
plotPlay = function(dfPlay, featureName=NULL, featureValue=NULL, runner=T) {
  dfRun = dfPlay %>% 
    filter(Position == "QB" | NflIdRusher == NflId)
  p = ggplot(dfPlay) +
    geom_point(aes(X, Y, color=Team)) +
    geom_vline(aes(xintercept=LineOfScrimmage), color="yellow") +
    geom_vline(aes(xintercept=10), color="grey") +
    geom_vline(aes(xintercept=110), color="grey") +
    coord_cartesian(xlim = c(0,120), ylim=c(0,53.3), expand=F) +
    theme(legend.title=element_blank()) +
    labs(title = paste("Offense", dfPlay$PossessionTeam, "; ",
      "PlayDirection", dfPlay$PlayDirection, "; ",
      "FieldPosition", dfPlay$FieldPosition, "; ",
      "HomeTeam", dfPlay$HomeTeamAbbr),
      subtitle = paste(featureName, round(featureValue,2), "; ", 
        "Yards", dfPlay$Yards, "; ",
        "Down", dfPlay$Down)) +
    geom_point(data=dfRun, aes(X, Y, color=Position))
  if (runner) plotRunner_add(p, dfPlay) else p
}
plotRunner_add = function(p, dfPlay) {
  dfRunner = dfPlay %>% 
    filter(NflIdRusher == NflId) %>% 
    mutate(
      dirRad = ((Dir + 90)%%360)/180,
      orRad = ((Orientation + 90)%%360)/180,
      Xd = X + S*cos(dirRad*pi),
      Yd = Y + S*sin(dirRad*pi),
      Xo = X + S/2*cos(orRad*pi),
      Yo = Y + S/2*sin(orRad*pi)
    )
  p +
    geom_segment(data=dfRunner, aes(x=X, y=Y, xend=Xd, yend=Yd, color="RB"), cex=1) +
    geom_segment(data=dfRunner, aes(x=X, y=Y, xend=Xo, yend=Yo, color="RB"), cex=1)
}
```

# Runner

Where is the runner? Where is the QB? What is the runner's position?

```{r}
dfPlay %>% 
  filter(Position == "QB")
dfPlay %>% 
  filter(NflIdRusher == NflId)
dfRun = dfPlay %>% 
  filter(Position == "QB" | NflIdRusher == NflId)
plotPlay(dfPlay) +
  geom_point(data=dfRun, aes(X, Y, color=Position))
```

Which way is runner facing? 
```{r}
plotRunner_add = function(p, dfPlay) {
  dfRunner = dfPlay %>% 
    filter(NflIdRusher == NflId) %>% 
    mutate(
      dirRad = ((Dir + 90)%%360)/180,
      orRad = ((Orientation + 90)%%360)/180,
      Xd = X + S*cos(dirRad*pi),
      Yd = Y + S*sin(dirRad*pi),
      Xo = X + S/2*cos(orRad*pi),
      Yo = Y + S/2*sin(orRad*pi)
    )
  p +
    geom_segment(data=dfRunner, aes(x=X, y=Y, xend=Xd, yend=Yd, color="RB"), cex=1) +
    geom_segment(data=dfRunner, aes(x=X, y=Y, xend=Xo, yend=Yo, color="RB"), cex=1)
}
dfRun = getRunnerDir(dfPlay)
dfRun = getRunnerDir(dfPlay, "Orientation")
p = plotPlay(dfPlay)
plotPlay(dfPlay) %>% 
  plotRunner_add(dfPlay)
```


