---
title: "Notes on "
author: "Kory D. Johnson"
date: 'Last Compiled `r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: show
    highlight: tango
    number_sections: yes
    theme: cosmo
    toc: yes
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE,
               collapse = FALSE,
               comment = "",
               strip.white = TRUE,
               warning = FALSE,
               message = FALSE,
               out.width = "70%",
               fig.align = "center")
```

## Topic

This file 

## Set Up  {.tabset .tabset-fade}

### Packages/Files {-}

```{r}
library(tidyverse)
library(magrittr)
```

### Load Data {-}

```{r, cache=T}
loadData <- function(fileName = "train.csv", ...) {
  file = paste("../input", fileName, sep="/")
  data.table::fread(file, ...) %>% 
    as_tibble()
}
df = loadData("trainClean.csv") %>% 
  mutate(YardsToGoal = YardLine + (100-2*YardLine)*(PossessionTeam==FieldPosition),
    LOS_Side = ifelse((PlayDirection=="right" & PossessionTeam==FieldPosition) |
        (PlayDirection=="left" & PossessionTeam!=FieldPosition),
      "left", "right"),
    LineOfScrimmage = ifelse(LOS_Side=="left", YardLine+10, 110-YardLine))
```

# Plotting Function

line of scrimmage isn't necessarily YardsToGoal, depnds on the side and direction of play.
If PlayDirection == right and in opponents field, then LOS is on right
If PlayDirection == right and in your field, then LOS is on Left
If PlayDirection == left and in opponents field, then LOS is on Left
If PlayDirection == leftand in your field, then LOS is on right.
Need to either add 10 or subtract 10 for endzone...

```{r}
dfPlay = filter(df, PlayId == df$PlayId[1])
plotPlay = function(dfPlay, featureName=NULL, featureValue=NULL, runner=T) {
  dfRun = dfPlay %>% 
    filter(Position == "QB" | NflIdRusher == NflId)
  p = ggplot(dfPlay) +
    geom_point(aes(X, Y, color=Team)) +
    geom_vline(aes(xintercept=LineOfScrimmage), color="yellow") +
    geom_vline(aes(xintercept=10), color="grey") +
    geom_vline(aes(xintercept=110), color="grey") +
    coord_cartesian(xlim = c(0,120), ylim=c(0,53.3), expand=F) +
    theme(legend.title=element_blank()) +
    labs(title = paste("Offense", dfPlay$PossessionTeam, "; ",
      "PlayDirection", dfPlay$PlayDirection, "; ",
      "FieldPosition", dfPlay$FieldPosition, "; ",
      "HomeTeam", dfPlay$HomeTeamAbbr),
      subtitle = paste(featureName, featureValue, "; ", 
        "Yards", dfPlay$Yards, "; ",
        "Down", dfPlay$Down)) +
    geom_point(data=dfRun, aes(X, Y, color=Position))
  if (runner) plotRunner_add(p, dfPlay) else p
}
plotRunner_add = function(p, dfPlay) {
  dfRunner = dfPlay %>% 
    filter(NflIdRusher == NflId) %>% 
    mutate(
      dirRad = ((-1^(PlayDirection=="left")*Dir + 90)%%360)/180,
      orRad = ((-1^(PlayDirection=="left")*Orientation + 90)%%360)/180,
      Xd = X + S*cos(dirRad*pi),
      Yd = Y + S*sin(dirRad*pi),
      Xo = X + S/2*cos(orRad*pi),
      Yo = Y + S/2*sin(orRad*pi)
    )
  p +
    geom_segment(data=dfRunner, aes(x=X, y=Y, xend=Xd, yend=Yd, color="RB"), cex=1) +
    geom_segment(data=dfRunner, aes(x=X, y=Y, xend=Xo, yend=Yo, color="RB"), cex=1)
}
plotPlay(dfPlay)
```

# Runner

Where is the runner? Where is the QB? What is the runner's position? Which way is runner facing? 


```{r}
dfPlay %>% 
  filter(Position == "QB")
dfPlay %>% 
  filter(NflIdRusher == NflId)
dfRun = dfPlay %>% 
  filter(Position == "QB" | NflIdRusher == NflId)
plotPlay(dfPlay) +
  geom_point(data=dfRun, aes(X, Y, color=Position))
```

# Looking at Plays

First, categorize runs: long, medium, short, bad
```{r}
ggplot(df) +
  geom_density(aes(Yards))
summary(df$Yards)
df %<>%
  mutate(YardsCategory = cut(Yards, breaks = c(-Inf, 0, 3, 6, Inf), 
             labels=c("bad", "short", "med", "long")))
table(df$YardsCategory)
PlayId_long = df %>% 
  filter(YardsCategory == "long") %>% 
  pull(PlayId)
PlayId_med = df %>% 
  filter(YardsCategory == "med") %>% 
  pull(PlayId)
PlayId_short = df %>% 
  filter(YardsCategory == "short") %>% 
  pull(PlayId)
PlayId_bad = df %>% 
  filter(YardsCategory == "bad") %>% 
  pull(PlayId)
```

Plot several runs from vector of ids
```{r}
plotPlays = function(ids, n=5) {
  for (index in sample.int(length(ids), 5)) {
    id = ids[index]
    p = df %>% 
      filter(PlayId == id) %>% 
      plotPlay()
    print(p)
  }
}
```

## Long Runs
```{r}
plotPlays(PlayId_long)
```

## Medium Runs
```{r}
plotPlays(PlayId_med)
```

## Short Runs
```{r}
plotPlays(PlayId_short)
```

## Bad Runs
```{r}
plotPlays(PlayId_bad)
```